---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
Load data set from zip file and store into data_set variable

```{r echo=TRUE}
data_set<-read.csv(unz("activity.zip", "activity.csv"),header = TRUE)
names(data_set)
head(data_set)
summary(data_set)

```

## What is mean total number of steps taken per day?
```{r echo=TRUE}
#Steps taken per day - Filter and aggregate data excluding null
sub_set <- aggregate(steps~date,data = data_set,FUN = sum,na.rm = T)
head(sub_set)

# Histogram
hist(sub_set$steps,xlab="Steps",ylab = "Frequency/Count",main="Total number of steps taken each day",col="green")

#mean
mean(sub_set$steps)

#median
median(sub_set$steps)


```


## What is the average daily activity pattern?
```{r echo=TRUE}
#Avg daily activity - Filter and aggregate data excluding null
avg_set <- aggregate(steps~interval,data = data_set,FUN = mean,na.rm = T)
head(avg_set)

plot(steps~interval,data=avg_set,type="l",xlab="Steps",ylab="Interval",col="red",main="Average daily activity")

#Interval maximum number of steps
avg_set$interval[which.max(avg_set$steps)]

```


## Imputing missing values
```{r echo=TRUE}

#Print missing steps
sum(is.na(data_set$steps)==TRUE)

# calculate mean for each interval
x<-aggregate(steps~interval,data = data_set,FUN = mean,na.rm = T)

#Create function replace null with average value
NA_Replace <- function(step,avgval){
  if(is.na(step)==TRUE){
    return(avgval)
  }
  return(step)
}

#new updated data set

data_set$steps<-mapply(NA_Replace,data_set$steps, x$steps)
head(data_set)


#Steps taken per day - Filter and aggregate data excluding null
sub_set <- aggregate(steps~date,data = data_set,FUN = sum,na.rm = T)
head(sub_set)

# Histogram
hist(sub_set$steps,xlab="Steps",ylab = "Frequency/Count",main="Total number of steps taken each day",col="green")

#mean
mean(sub_set$steps)

#median
median(sub_set$steps)

```
**Answer To Question:**
The mean and the median are now almost the same after replacing missing values with the mean value for the relevant interval. It makes sense that the median value would now move closer to the mean. So the Median value increased after this method of missing value replacement.




## Are there differences in activity patterns between weekdays and weekends?
